Experimentally Informed Site-Specific Substitution Models Substantially Deepen Viral Divergence Estimates

### 2018-01-18 manuscript/manuscript.tex
rough outlines of last two sections

\subsection{Competing effects of shifting preferences and long branches.}

@@

>>>>

\begin{itemize}
\item We tested the effect substitution model choice on branch length estimation using influenza HA sequences. 
\begin{itemize}
\item These sequences presumably evolve with site-specific frequencies but unlike the simulations we don't \textit{know} the model
\item These sequences are from fairly diverged proteins. Some sequences share only 40\% identity on the amino acid level 
\item HA has small, closely related subtypes with long branches between subtypes and groups. 
\end{itemize}
\item Both $\Gamma\omega$ and non-uniform stationary states result in an extension of branch lengths. 
\begin{itemize}
\item The addition of $\Gamma\omega$ extends branches independent of preferences. 
\item The addition of preferences extends branches \textit{from the focal sequence} independent of $\Gamma\omega$. 
\end{itemize}
\item The ExpCM does not have a uniform effect across the whole tree. 
\begin{itemize}
\item This is not too surprising. DMS makes accurate measurements in a given genetic background. 
\item You would expect epistatic effects over such a large divergence distance. ``Shifting preferences"
\end{itemize}
\end{itemize}

\subsection*{Competing effects of shifting preferences and long branches.}

@@

\begin{itemize}
\item We investigated the competing effects of shifting preferences and long branch estimation. That is, our preferences are most relevant to sequences they are close to but the effect of a non-uniform stationary is only observed on long branches. 
\begin{itemize}
\item We broke the large tree into small, sub trees with differing maximum divergence levels. 
\item We estimated the branch lengths using the same models we used for empirical data above. 
\end{itemize}
\item The ``relevance" of the preferences is inversely correlated with sequence divergence. 
\begin{itemize}
\item We used the ExpCM stringency parameter as the measure of preference ``relevance". The larger the $\beta$ value the more ``relevant" the preferences. 
\item As the maximum divergence of the tree increases the $\beta$ value decreases. 
\end{itemize}
\item All models estimate short branches relatively equally.

\end{itemize}
